---
title: 'La Liga'
author: "Group 12"
date: "November 21, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description of "La Liga" (Spain soccer)
"La Liga" is the professional soccer league of Spain. The League follows a double round competition format in which each team meets all other contestants (as home team and away team). Each season, La Liga has 20 teams. Thus, there are 10 games each week in a tournament of 38 weeks (match days). Hence, each season has a total of 380 games. Each team receives 3 points for a win, 1 point for a draw, and 0 points for a loss. Teams are ranked by total points, and the team ranked first at the end of the season is the Champion. 
# Data set
The data set can be accessed online here: 
https://datahub.io/sports-data/spanish-la-liga#resource-season-1819 
There are 10 csv files that includes the results for each game of the current season and the last 9 seasons.
Some of the columns include the following information:
Div = League Division
Date = Match Date (dd/mm/yy)
HomeTeam = Home Team
AwayTeam = Away Team
FTHG= Full Time Home Team Goals
FTAG= Full Time Away Team Goals
FTR= Full Time Result (H=Home Win, D=Draw, A=Away Win)
HTHG = Half Time Home Team Goals
HTAG = Half Time Away Team Goals
HTR = Half Time Result (H=Home Win, D=Draw, A=Away Win)
HS = Home Team Shots
AS = Away Team Shots
HST = Home Team Shots on Target
AST = Away Team Shots on Target
HC = Home Team Corners
AC = Away Team Corners
HF = Home Team Fouls Committed
AF = Away Team Fouls Committed
HY = Home Team Yellow Cards
AY = Away Team Yellow Cards
HR = Home Team Red Cards
AR = Away Team Red Cards
A link for the complete description of the columns is here:  http://www.football-data.co.uk/notes.txt

# Getting started
### Here we get a csv file for all the seasons and the 22 columns we are interested in
```{r}
getwd() #gives the working directory
setwd("C:/Users/mserrano/Documents/STATS/Stat 579/Final Project/LaLiga") # note you have to use this slash "/" if you are using a PC. Be sure all your csv files are in this directory.
season0910<- read.csv("season-0910.csv")
season1011<- read.csv("season-1011.csv")
season1112<- read.csv("season-1112.csv")
season1213<- read.csv("season-1213.csv")
season1314<- read.csv("season-1314.csv")
season1415<- read.csv("season-1415.csv")
season1516<- read.csv("season-1516.csv")
season1617<- read.csv("season-1617.csv")
season1718<- read.csv("season-1718.csv")

#selecting the same number columns (22) for each season
library(dplyr)
season0910 <- season0910 %>% select(Div,Date,HomeTeam,AwayTeam,FTHG,FTAG,FTR,HTHG,HTAG,HTR,HS,AS,HST,AST,HF,AF,HC,AC,HY,                      AY,HR,AR) 
season1011 <- season1011 %>% select(Div,Date,HomeTeam,AwayTeam,FTHG,FTAG,FTR,HTHG,HTAG,HTR,HS,AS,HST,AST,HF,AF,HC,AC,HY,                      AY,HR,AR) 
season1112 <- season1112 %>% select(Div,Date,HomeTeam,AwayTeam,FTHG,FTAG,FTR,HTHG,HTAG,HTR,HS,AS,HST,AST,HF,AF,HC,AC,HY,                      AY,HR,AR) 
season1213 <- season1213 %>% select(Div,Date,HomeTeam,AwayTeam,FTHG,FTAG,FTR,HTHG,HTAG,HTR,HS,AS,HST,AST,HF,AF,HC,AC,HY,                      AY,HR,AR) 
season1314 <- season1314 %>% select(Div,Date,HomeTeam,AwayTeam,FTHG,FTAG,FTR,HTHG,HTAG,HTR,HS,AS,HST,AST,HF,AF,HC,AC,HY,                      AY,HR,AR) 
season1415 <- season1415 %>% select(Div,Date,HomeTeam,AwayTeam,FTHG,FTAG,FTR,HTHG,HTAG,HTR,HS,AS,HST,AST,HF,AF,HC,AC,HY,                      AY,HR,AR) 
season1516 <- season1516 %>% select(Div,Date,HomeTeam,AwayTeam,FTHG,FTAG,FTR,HTHG,HTAG,HTR,HS,AS,HST,AST,HF,AF,HC,AC,HY,                      AY,HR,AR) 
season1617 <- season1617 %>% select(Div,Date,HomeTeam,AwayTeam,FTHG,FTAG,FTR,HTHG,HTAG,HTR,HS,AS,HST,AST,HF,AF,HC,AC,HY,                      AY,HR,AR) 
season1718 <- season1718 %>% select(Div,Date,HomeTeam,AwayTeam,FTHG,FTAG,FTR,HTHG,HTAG,HTR,HS,AS,HST,AST,HF,AF,HC,AC,HY,                      AY,HR,AR) 

allseasons <- rbind(season0910,season1011,season1112,season1213,season1314,season1415,
                    season1516,season1617, season1718)
write.csv(allseasons,"allseasons.csv")
```


# Part5
5- When a team get a red card, a) does it imply that the team is going to lose the game? b) In the cases a team which is shown a red card loses the game, is the goal difference larger than the goal difference in  games  without red card? 
For this part of the exploration we will select those games where a red card was shown to the team that lost the game (the red card implies a player is kicked out of the game). Then we will calculate the goal difference in those games.
From these variable we formulate the following hypothesis:
.	We hypothesize that the team with a red card has higher chances to lose the game.
.	Also, the goal difference in those games when a red card is shown is larger in comparison to games with no red cards.

```{r}
head(allseasons)
#create a column for red_win  and red_lose
#allseasons<- allseasons %>% mutate(red_win=ifelse(HR>0 &FTR=="H"| AR>0 &FTR=="A",1,0),
#                      red_lose=ifelse(HR>0 &FTR=="A" | AR>0 &FTR=="H",1,0))

allseasons<- allseasons %>% mutate(red_WL=ifelse(HR>0 &FTR=="H"| AR>0 &FTR=="A","red_win",
                                                (ifelse(HR>0 &FTR=="A" | AR>0 &FTR=="H","red_lose",0))))

#create a graph using ggplot. 
#red_win: got red card and won the game
#red_lose: got red card and lose the game
library(ggplot2)                                                                                          
allseasons %>% filter(red_WL %in% c("red_win","red_lose")) %>% 
  ggplot(aes(x=red_WL,fill=red_WL)) + geom_bar()

#create a column for the goal difference. Note the use of abs() for absolute value.
#create a column to indicate a red card was shown on either home team or away team
allseasons<- allseasons %>% mutate(goal_dif=abs(FTHG - FTAG),
                                   red_shown=ifelse(HR>0|AR>0,"Red_shown","No_Red_cards"))

#Graph for average goal diference per game for "Red_shown" and "No_Red_cards"
allseasons %>% group_by(red_shown) %>% summarize(mean_goal_dif=mean(goal_dif)) 
allseasons %>% group_by(red_shown) %>% summarize(mean_goal_dif=mean(goal_dif)) %>%
  ggplot(aes(x=red_shown, y=mean_goal_dif, fill= red_shown)) + geom_col()

#Graph for average goal difference per game for Red_lose and Red_win
allseasons %>% group_by(red_WL) %>% summarize(mean_goal_dif=mean(goal_dif)) 
allseasons %>% group_by(red_WL) %>% summarize(mean_goal_dif=mean(goal_dif)) %>%
  ggplot(aes(x=red_WL, y=mean_goal_dif, fill=red_WL)) + geom_col()


```



